---
title: "Bird Strikes on AirCraft Analysis <br> <span style='font-size: 18px; color:
  grey;'>by Group 5</span>"
author:
- Mingxi Li (li.mingxi1@northeastern.edu)
- Si Wu (wu.si3@northeastern.edu)
- Zheng Zheng (zheng.zheng2@northeastern.edu)
- Jiawei Zhou (zhou.jiaw@northeastern.edu)
output:
  pdf_document:
    toc: yes
  html_notebook:
    theme: cerulean
    toc: yes
    toc_float: yes
    number_sections: yes
  html_document:
    toc: yes
    df_print: paged
---

## Introduction

Bird strikes have been a matter of concern for the aviation industry since its inception. These events, which refer to the collision of a bird with an aircraft, can result in substantial financial losses and potential risks to safety. As the aviation industry grows, understanding and mitigating bird strikes becomes more crucial, especially considering the rise in global air traffic.

In this project, our group - Group 5, dives deep into the data related to bird strikes.Our primary aim is to analyze the patterns of bird strikes on aircrafts across different years, especially focusing on various flight phases. By doing so, we hope to identify high-risk factors and periods associated with these incidents."

With this analysis, we hope to contribute valuable insights that can assist airlines, pilots, and airport authorities in understanding the dynamics of bird strikes better. Our findings might pave the way for innovative solutions to tackle this persistent challenge in the aviation industry.

Source data: FAA[1] The data for our analysis is sourced from a comprehensive database that logs bird strikes on aircrafts. This database provides detailed insights on the various factors and conditions under which these incidents occur, making it a rich source for our study

## Build ERD Model and Relational Schema

<div>

1.  The conceptual model as a Crow's Foot ERD - Zheng Zheng

    ![](Conceptual%20Model%20(crow's%20foot).png)

2.  The logical model as a Crow's Feet Relational Schema - Zheng Zheng

    ![](Database%20ER%20diagram%20(crow's%20foot).png)

</div>

## Create the database

###Connect to MySQL server (Remember to change the password for your own laptop)

```{r}
# Load the required library
library(RMySQL)

# Settings
host <- "localhost"
port <- 3306
user <- "root"
password <- ""

# Create a connection
con <- dbConnect(MySQL(), user='root', 
                 password='Zz19920608!', host='localhost')

# Create the database
dbGetQuery(con, "DROP DATABASE IF EXISTS BirdStrikesOnAircraft")
dbGetQuery(con, "CREATE DATABASE BirdStrikesOnAircraft")

# Use the created database
dbGetQuery(con, "USE BirdStrikesOnAircraft")

```

### Create the tables

```{r}
# 9. Airline/Operator
dbGetQuery(con, "DROP TABLE IF EXISTS Airline_Operator")
dbGetQuery(con, "CREATE TABLE Airline_Operator(
    Airline_OperatorID INTEGER AUTO_INCREMENT PRIMARY KEY,
    Airline_OperatorName VARCHAR(255)
)")

# 1. Aircraft
dbGetQuery(con, "DROP TABLE IF EXISTS Aircraft")
dbGetQuery(con, "CREATE TABLE Aircraft(
    AircraftID INTEGER AUTO_INCREMENT PRIMARY KEY,
    Type VARCHAR(255),
    Make_Model VARCHAR(255),
    Number_of_Engines INTEGER,
    Airline_OperatorID INTEGER,
    IsLarge VARCHAR(255),
    FOREIGN KEY (Airline_OperatorID) REFERENCES Airline_Operator(Airline_OperatorID)
)")

# 2. Airport
dbGetQuery(con, "DROP TABLE IF EXISTS Airport")
dbGetQuery(con, "CREATE TABLE Airport(
    AirportID INTEGER AUTO_INCREMENT PRIMARY KEY,
    AirportName VARCHAR(255)
)")

# 5. SkyConditions
dbGetQuery(con, "DROP TABLE IF EXISTS SkyConditions")
dbGetQuery(con, "CREATE TABLE SkyConditions(
    SkyConditionID INTEGER AUTO_INCREMENT PRIMARY KEY,
    SkyConditionDescription VARCHAR(255)
)")

# 4. FlightPhase
dbGetQuery(con, "DROP TABLE IF EXISTS FlightPhase")
dbGetQuery(con, "CREATE TABLE FlightPhase(
    PhaseID INTEGER AUTO_INCREMENT PRIMARY KEY,
    FlightPhaseDescription VARCHAR(255)
)")

# 6. Wildlife
dbGetQuery(con, "DROP TABLE IF EXISTS Wildlife")
dbGetQuery(con, "CREATE TABLE Wildlife(
    WildlifeID INTEGER AUTO_INCREMENT PRIMARY KEY,
    WildlifeSpecies VARCHAR(255),
    WildlifeSize VARCHAR(50)
)")

# 8. Effect
dbGetQuery(con, "DROP TABLE IF EXISTS Effect")
dbGetQuery(con, "CREATE TABLE Effect(
    EffectID INTEGER AUTO_INCREMENT PRIMARY KEY,
    EffectDescription VARCHAR(255)
)")

# 3. Flight
dbGetQuery(con, "DROP TABLE IF EXISTS Flight")
dbGetQuery(con, "CREATE TABLE Flight(
    FlightID INTEGER AUTO_INCREMENT PRIMARY KEY,
    FlightDate DATE,
    OriginState VARCHAR(255),
    AircraftID INTEGER,
    AirportID INTEGER,
    FOREIGN KEY (AircraftID) REFERENCES Aircraft(AircraftID),
    FOREIGN KEY (AirportID) REFERENCES Airport(AirportID)
)")

# 7. BirdStrikeEvent
dbGetQuery(con, "DROP TABLE IF EXISTS BirdStrikeEvent")
dbGetQuery(con, "CREATE TABLE BirdStrikeEvent(
    RecordID INTEGER AUTO_INCREMENT PRIMARY KEY,
    WildlifeID INTEGER,
    FlightID INTEGER,
    PhaseID INTEGER,
    SkyConditionID INTEGER,
    EffectID INTEGER,
    IndicatedDamage BOOLEAN,
    AltitudeBin VARCHAR(100),
    FeetAboveGround INTEGER,
    NumberOfPeopleInjured INTEGER,
    Costs DECIMAL(12,2),
    PilotWarned VARCHAR(100),
    RemainsCollected VARCHAR(100),
    RemainsSent VARCHAR(100),
    Remark TEXT,
    NumberStruckRange VARCHAR(100),
    NumberStruckActual INTEGER,
    FOREIGN KEY (WildlifeID) REFERENCES Wildlife(WildlifeID),
    FOREIGN KEY (FlightID) REFERENCES Flight(FlightID),
    FOREIGN KEY (PhaseID) REFERENCES FlightPhase(PhaseID),
    FOREIGN KEY (SkyConditionID) REFERENCES SkyConditions(SkyConditionID),
    FOREIGN KEY (EffectID) REFERENCES Effect(EffectID)
)")

```

## Load the Bird Strikes CSV file into R and populate the tables with the data.

```{r}
library(DBI)

# Load data from CSV
csv_data <- read.csv('BirdStrikesData.csv')
```

```{r}
colnames(csv_data)
```

```{r}
names(csv_data) <- c(
  "RecordID", 
  "Type", 
  "AirportName", 
  "AltitudeBin", 
  "Make_Model", 
  "NumberStruckRange", 
  "NumberStruckActual", 
  "EffectDescription", 
  "FlightDate", 
  "IndicatedDamage", 
  "NumberOfEngines", 
  "Airline_OperatorName", 
  "OriginState", 
  "FlightPhaseDescription", 
  "Precipitation", 
  "RemainsCollected", 
  "RemainsSent",
  "Remarks",
  "WildlifeSize",
  "SkyConditionDescription",
  "WildlifeSpecies",
  "PilotWarned",
  "Costs",
  "FeetAboveGround",
  "NumberOfPeopleInjured",
  "IsLarge"
)

```

```{r}
# Aircraft Table
csv_data$Type <- as.character(csv_data$Type)
csv_data$Make_Model <- as.character(csv_data$Make_Model)
csv_data$NumberOfEngines <- as.integer(as.character(csv_data$NumberOfEngines))
csv_data$IsLarge <- as.character(csv_data$IsLarge)

# Airport Table
csv_data$AirportName <- as.character(csv_data$AirportName)

# Flight Table
csv_data$FlightDate <- as.Date(csv_data$FlightDate, format="%m/%d/%Y")
csv_data$OriginState <- as.character(csv_data$OriginState)

# FlightPhase Table
csv_data$FlightPhaseDescription <- as.character(csv_data$FlightPhaseDescription)

# SkyConditions Table
csv_data$SkyConditionDescription <- as.character(csv_data$SkyConditionDescription)

# Wildlife Table
csv_data$WildlifeSpecies <- as.character(csv_data$WildlifeSpecies)
csv_data$WildlifeSize <- as.character(csv_data$WildlifeSize)

# BirdStrikeEvent Table
csv_data$AltitudeBin <- as.character(csv_data$AltitudeBin)
csv_data$FeetAboveGround <- as.integer(as.character(csv_data$FeetAboveGround))
csv_data$NumberOfPeopleInjured <- as.integer(as.character(csv_data$NumberOfPeopleInjured))
csv_data$Costs <- as.numeric(as.character(csv_data$Costs))
csv_data$PilotWarned <- as.character(csv_data$PilotWarned)
csv_data$RemainsCollected <- as.character(csv_data$RemainsCollected)
csv_data$RemainsSent <- as.character(csv_data$RemainsSent)
csv_data$Remarks <- as.character(csv_data$Remarks)
csv_data$NumberStruckRange <- as.character(csv_data$NumberStruckRange)
csv_data$NumberStruckActual <- as.integer(as.character(csv_data$NumberStruckActual))
csv_data$IndicatedDamage <- ifelse(csv_data$IndicatedDamage == 'Caused damage', 1, 0)

# Effect Table
csv_data$EffectDescription <- as.character(csv_data$EffectDescription)

# Airline/Operator Table
csv_data$Airline_OperatorName <- as.character(csv_data$Airline_OperatorName)

```

```{r}
# Identify columns with categorical data 
categorical_cols <- c(
  "Type", "AirportName", "AltitudeBin", "Make_Model", "EffectDescription", 
  "OriginState", "FlightPhaseDescription", "Precipitation", "RemainsCollected", 
  "RemainsSent", "Remarks", "WildlifeSize", "SkyConditionDescription", "WildlifeSpecies", 
  "PilotWarned", "IsLarge", "Airline_OperatorName", "NumberStruckRange"
)

# Replace NA values with 'Unknown' for these columns
csv_data[categorical_cols] <- lapply(csv_data[categorical_cols], function(col) {
    ifelse(is.na(col)| col == "", 'Unknown', col)
})

# Identify columns with numerical data
numerical_cols <- c(
  "RecordID", "NumberStruckActual", "NumberOfEngines", "Costs", 
  "FeetAboveGround", "NumberOfPeopleInjured"
)

# Replace 'Unknown' or other non-date values in FlightDate with '1900-01-01'
csv_data$FlightDate[csv_data$FlightDate == ""  | !grepl("^\\d{4}-\\d{2}-\\d{2}$", csv_data$FlightDate)] <- '1900-01-01'

# Replace NA values with 0 for these columns
csv_data[numerical_cols] <- lapply(csv_data[numerical_cols], function(col) {
    ifelse(is.na(col)| col == "", 0, col)
})

```

```{r}
# Airline/Operator IDs
airline_operator_ids <- unique(csv_data$Airline_OperatorName)
airline_operator_ids <- data.frame(Airline_OperatorID = seq_along(airline_operator_ids),
                                   Airline_OperatorName = airline_operator_ids)

# Aircraft IDs
aircraft_data <- unique(csv_data[, c("Type", "Make_Model", "NumberOfEngines", "Airline_OperatorName", "IsLarge")])
aircraft_data <- merge(aircraft_data, airline_operator_ids, by="Airline_OperatorName")
aircraft_data$AircraftID <- seq_along(aircraft_data$Type)

# Airport IDs
airport_ids <- unique(csv_data$AirportName)
airport_ids <- data.frame(AirportID = seq_along(airport_ids), AirportName = airport_ids)

# FlightPhase IDs
flight_phase_ids <- unique(csv_data$FlightPhaseDescription)
flight_phase_ids <- data.frame(PhaseID = seq_along(flight_phase_ids), FlightPhaseDescription = flight_phase_ids)

# SkyConditions IDs
sky_condition_ids <- unique(csv_data$SkyConditionDescription)
sky_condition_ids <- data.frame(SkyConditionID = seq_along(sky_condition_ids), SkyConditionDescription = sky_condition_ids)

# Wildlife IDs
wildlife_data <- unique(csv_data[, c("WildlifeSpecies", "WildlifeSize")])
wildlife_data$WildlifeID <- seq_along(wildlife_data$WildlifeSpecies)

# Effect IDs
effect_ids <- unique(csv_data$EffectDescription)
effect_ids <- data.frame(EffectID = seq_along(effect_ids), EffectDescription = effect_ids)

# Generating FlightID for each row in the csv_data dataframe
csv_data$FlightID <- seq(1, nrow(csv_data))

```

```{r}
# Merge Airline/Operator IDs
csv_data <- merge(csv_data, airline_operator_ids, by="Airline_OperatorName", all.x = TRUE)

# Merge Aircraft IDs
csv_data <- merge(csv_data, aircraft_data[, c("Type", "Make_Model", "NumberOfEngines", "Airline_OperatorID", "IsLarge", "AircraftID")],
                 by = c("Type", "Make_Model", "NumberOfEngines", "Airline_OperatorID", "IsLarge"), all.x = TRUE)

# Merge Airport IDs
csv_data <- merge(csv_data, airport_ids, by="AirportName", all.x = TRUE)

# Merge FlightPhase IDs
csv_data <- merge(csv_data, flight_phase_ids, by="FlightPhaseDescription", all.x = TRUE)

# Merge SkyConditions IDs
csv_data <- merge(csv_data, sky_condition_ids, by="SkyConditionDescription", all.x = TRUE)

# Merge Wildlife IDs
csv_data <- merge(csv_data, wildlife_data, by = c("WildlifeSpecies", "WildlifeSize"), all.x = TRUE)

# Merge Effect IDs
csv_data <- merge(csv_data, effect_ids, by="EffectDescription", all.x = TRUE)

```

```{r}
# Clear data from tables
dbGetQuery(con, "DELETE FROM BirdStrikeEvent")
dbGetQuery(con, "DELETE FROM Flight")
dbGetQuery(con, "DELETE FROM Wildlife")
dbGetQuery(con, "DELETE FROM SkyConditions")
dbGetQuery(con, "DELETE FROM FlightPhase")
dbGetQuery(con, "DELETE FROM Effect")
dbGetQuery(con, "DELETE FROM Aircraft")
dbGetQuery(con, "DELETE FROM Airport")
dbGetQuery(con, "DELETE FROM Airline_Operator")

```

```{r}
#1. Airline/Operator
for(i in 1:nrow(airline_operator_ids)) {
  query <- sprintf("INSERT INTO Airline_Operator (Airline_OperatorID, Airline_OperatorName) VALUES (%d, '%s')",
                   airline_operator_ids$Airline_OperatorID[i], airline_operator_ids$Airline_OperatorName[i])
  dbGetQuery(con, query)
}

```

```{r}
# Insert data into Airline/Operator
for(i in 1:nrow(aircraft_data)) {
  query <- sprintf("INSERT INTO Aircraft (AircraftID, Type, Make_Model, Number_of_Engines, Airline_OperatorID, IsLarge) VALUES (%d, '%s', '%s', %d, %d, '%s')",
                   aircraft_data$AircraftID[i], aircraft_data$Type[i], aircraft_data$Make_Model[i], aircraft_data$NumberOfEngines[i],
                   aircraft_data$Airline_OperatorID[i], aircraft_data$IsLarge[i])
  dbGetQuery(con, query)
}


```

```{r}
# Insert data into Airport
dbGetQuery(con, "DELETE FROM Airport")
escape_quotes <- function(string) {
  return(gsub("'", "''", string))
}

# Insert data into Airport
for(i in 1:nrow(airport_ids)) {
  airport_name_escaped <- escape_quotes(airport_ids$AirportName[i])
  query <- sprintf("INSERT INTO Airport (AirportID, AirportName) VALUES (%d, '%s')", 
                   airport_ids$AirportID[i], airport_name_escaped)
  dbGetQuery(con, query)
}
```

```{r}
# 4. FlightPhase Table
for(i in 1:nrow(flight_phase_ids)) {
  query <- sprintf("INSERT INTO FlightPhase (PhaseID, FlightPhaseDescription) VALUES (%d, '%s')", 
                   flight_phase_ids$PhaseID[i], flight_phase_ids$FlightPhaseDescription[i])
  dbGetQuery(con, query)
}

```

```{r}
# 5. SkyConditions Table
for(i in 1:nrow(sky_condition_ids)) {
  query <- sprintf("INSERT INTO SkyConditions (SkyConditionID, SkyConditionDescription) VALUES (%d, '%s')", 
                   sky_condition_ids$SkyConditionID[i], sky_condition_ids$SkyConditionDescription[i])
  dbGetQuery(con, query)
}

```

```{r}
#6. Wildlife Table
escape_quotes <- function(string) {
  return(gsub("'", "''", string))
}

for(i in 1:nrow(wildlife_data)) {
  species_escaped <- escape_quotes(wildlife_data$WildlifeSpecies[i])
  size_escaped <- escape_quotes(wildlife_data$WildlifeSize[i])

  query <- sprintf("INSERT INTO Wildlife (WildlifeID, WildlifeSpecies, WildlifeSize) VALUES (%d, '%s', '%s')", 
                   wildlife_data$WildlifeID[i], species_escaped, size_escaped)
  dbGetQuery(con, query)
}

```

```{r}


# Insert data into Flight table
for(i in 1:nrow(csv_data)) {
  query <- sprintf("INSERT INTO Flight (FlightID, FlightDate, OriginState, AircraftID, AirportID) VALUES (%d, '%s', '%s', %d, %d)", 
                   csv_data$FlightID[i], csv_data$FlightDate[i], escape_quotes(csv_data$OriginState[i]), csv_data$AircraftID[i], csv_data$AirportID[i])
  dbGetQuery(con, query)
}
```

```{r}
#Insert the data into the Effect table
for(i in 1:nrow(effect_ids)) {
    query <- sprintf("INSERT INTO Effect (EffectID, EffectDescription) VALUES (%d, '%s')", 
                     effect_ids$EffectID[i], effect_ids$EffectDescription[i])
    dbGetQuery(con, query)
}
```

```{r}
# Insert data into BirdStrikeEvent
sanitize_string <- function(input_string) {
  # Replace single quotes with two single quotes for SQL
  output_string <- gsub("'", "''", input_string)
  
  # Add more replacements as needed
  
  return(output_string)
}

# Apply the function to character columns in your dataframe
char_cols <- sapply(csv_data, is.character)
csv_data[char_cols] <- lapply(csv_data[char_cols], sanitize_string)

dbGetQuery(con, "DELETE FROM BirdStrikeEvent")
for(i in 1:nrow(csv_data)) {
  query <- sprintf("INSERT INTO BirdStrikeEvent (RecordID, WildlifeID, FlightID, PhaseID, SkyConditionID, EffectID, IndicatedDamage, AltitudeBin, FeetAboveGround, NumberOfPeopleInjured, Costs, PilotWarned, RemainsCollected, RemainsSent, Remark, NumberStruckRange, NumberStruckActual) VALUES (%d, %d, %d, %d, %d, %d, %d, '%s', %d, %d, %.2f, '%s', '%s', '%s', '%s', '%s', %d)", 
                   csv_data$RecordID[i], csv_data$WildlifeID[i], csv_data$FlightID[i], csv_data$PhaseID[i], csv_data$SkyConditionID[i], csv_data$EffectID[i], csv_data$IndicatedDamage[i], csv_data$AltitudeBin[i], csv_data$FeetAboveGround[i], csv_data$NumberOfPeopleInjured[i], csv_data$Costs[i], csv_data$PilotWarned[i], csv_data$RemainsCollected[i], csv_data$RemainsSent[i], csv_data$Remarks[i], csv_data$NumberStruckRange[i], csv_data$NumberStruckActual[i])
  dbGetQuery(con, query)
}


```

## Analysis Tasks - Si Wu and Jiawei Zhou

# Create a SQL query against your database to find the number of bird strike incidents for each airline upon take-off or climb. Include all airlines. You may either use a {sql} code chunk or an R function to execute the query. It must be a single query.

```{sql connection=con}
SELECT 
    ao.Airline_OperatorName AS Airline,
    COUNT(*) AS NumberOfIncidents
FROM 
    BirdStrikeEvent bse
JOIN 
    Flight f ON bse.FlightID = f.FlightID
JOIN 
    Aircraft ac ON f.AircraftID = ac.AircraftID
JOIN 
    Airline_Operator ao ON ac.Airline_OperatorID = ao.Airline_OperatorID
JOIN 
    FlightPhase fp ON bse.PhaseID = fp.PhaseID
WHERE 
    fp.FlightPhaseDescription IN ('Take-Off', 'Climb')
GROUP BY 
    ao.Airline_OperatorName
ORDER BY 
    NumberOfIncidents DESC;

```

```{sql connection=con}

SELECT AO.Airline_OperatorName AS Airline,
       SUM(CASE WHEN BE.PhaseID IN (1, 2) THEN 1 ELSE 0 END) AS BirdStrikeCount
FROM Airline_Operator AS AO
LEFT JOIN Aircraft AS A ON AO.Airline_OperatorID = A.Airline_OperatorID
LEFT JOIN Flight AS F ON A.AircraftID = F.AircraftID
LEFT JOIN BirdStrikeEvent AS BE ON F.FlightID = BE.FlightID
GROUP BY AO.Airline_OperatorName
ORDER BY BirdStrikeCount DESC;
```

```{sql connection=con}
SELECT * 
FROM FlightPhase
;
```

#### Create a SQL query against your database to find the airports that had the most bird strike incidents (during any flight phase). Include all airlines. You may either use a *{sql}* code chunk or an R function to execute the query. It must be a single query.

```{sql connection=con}
SELECT 
    ap.AirportName AS Airport,
    COUNT(*) AS NumberOfIncidents
FROM 
    BirdStrikeEvent bse
JOIN 
    Flight f ON bse.FlightID = f.FlightID
JOIN 
    Airport ap ON f.AirportID = ap.AirportID
GROUP BY 
    ap.AirportName
ORDER BY 
    NumberOfIncidents DESC;

```

#### Create a SQL query against your database to find the number of bird strike incidents by year. Include all airlines. You may either use a *{sql}* code chunk or an R function to execute the query. It must be a single query.

```{sql connection=con}
SELECT 
    YEAR(f.FlightDate) AS Year,
    COUNT(*) AS NumberOfStrikes
FROM 
    Flight f
JOIN 
    Aircraft a ON f.AircraftID = a.AircraftID
JOIN 
    Airline_Operator ao ON a.Airline_OperatorID = ao.Airline_OperatorID
JOIN 
    BirdStrikeEvent bse ON f.FlightID = bse.FlightID

WHERE 
    YEAR(f.FlightDate) <> 1900
GROUP BY 
    Year
ORDER BY 
    Year;

```

# Using the above data, build a column chart that visualizes the number of bird strikes incidents per year from 2008 to 2011 during take-off/climbing and during descent/approach/landing, so you can compare one versus the other. Adorn the graph with appropriate axis labels, titles, legend, data labels, etc.

```{sql connection=con}
SELECT 
    YEAR(f.FlightDate) AS Year,
    fp.FlightPhaseDescription,
    COUNT(*) AS NumberOfIncidents
FROM 
    BirdStrikeEvent bse
JOIN 
    Flight f ON bse.FlightID = f.FlightID
JOIN 
    FlightPhase fp ON bse.PhaseID = fp.PhaseID
WHERE 
    YEAR(f.FlightDate) BETWEEN 2008 AND 2011
    AND fp.FlightPhaseDescription IN ('Take-off', 'Climbing', 'Descent', 'Approach', 'Landing')
GROUP BY 
    YEAR(f.FlightDate),
    fp.FlightPhaseDescription
HAVING 
    fp.FlightPhaseDescription IN ('Take-off', 'Climbing', 'Descent', 'Approach', 'Landing')
ORDER BY 
    YEAR(F.FlightDate), FP.FlightPhaseDescription;
```

```{r}
library(ggplot2)
library(DBI)

query <- "
SELECT 
    YEAR(f.FlightDate) AS Year,
    fp.FlightPhaseDescription,
    COUNT(*) AS NumberOfIncidents
FROM 
    BirdStrikeEvent bse
JOIN 
    Flight f ON bse.FlightID = f.FlightID
JOIN 
    FlightPhase fp ON bse.PhaseID = fp.PhaseID
WHERE 
    YEAR(f.FlightDate) BETWEEN 2008 AND 2011
    AND fp.FlightPhaseDescription IN ('Take-off', 'Climbing', 'Descent', 'Approach', 'Landing')
GROUP BY 
    YEAR(f.FlightDate),
    fp.FlightPhaseDescription
HAVING 
    fp.FlightPhaseDescription IN ('Take-off', 'Climbing', 'Descent', 'Approach', 'Landing')
ORDER BY 
    YEAR(F.FlightDate), FP.FlightPhaseDescription;
"

data <- dbGetQuery(con, query)

ggplot(data, aes(x=factor(Year), y=NumberOfIncidents, fill=FlightPhaseDescription)) + 
  geom_bar(stat="identity", position="dodge") +
  labs(title="Bird Strike Incidents per Year (2008-2011)", 
       x="Year", 
       y="Number of Incidents") +
  scale_fill_brewer(palette="Set2") +
  theme_minimal()


```

#### Create a stored procedure in MySQL (note that if you used SQLite, then you cannot complete this step) that removes a bird strike incident from the database. You may decide what you need to pass to the stored procedure to remove a bird strike incident, *e.g.*, departure airport, airlines, or some ID. Show that the deletion worked as expected.

List a RecordID to future verification

```{sql connection=con}
SELECT * FROM BirdStrikeEvent WHERE RecordID = 1195;
```

Create a procedure table

```{sql connection=con}
CREATE PROCEDURE DeleteBirdStrikeEvent(IN incidentRecordID INT)
BEGIN
  DELETE FROM BirdStrikeEvent WHERE RecordID = incidentRecordID;
END 
```

To invoke the procedure and remove a bird strike event with, for example, RecordID = 5:

```{sql connection=con}
CALL DeleteBirdStrikeEvent(1195);
```

To verify the deletion:

```{sql connection=con}
SELECT * FROM BirdStrikeEvent WHERE RecordID = 1195;
```

```{r}
# Close the connection
dbDisconnect(con)
```
